---
title: "cmip6_warming_levels"
author: "Doug McNeall"
date: "8/4/2020"
output: html_document
---

Extract year of passing warming levels and co2 concetration at the time in the cmip6 archive.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

# Load packages
library(ncdf4)
library(fields)

# r = realization index
# i = initialization index
# p = physics index
# f = forcing index
```



### Helper functions
```{r}
extractModelnames = function(filenames, startstring, stopstring){
  # Function to extract model names from a vector of cmip filenames
  # that have similar startstring and stopstring

  fs = lapply(filenames,regexpr, pattern = startstring)
  fe = lapply(filenames,regexpr, pattern = stopstring)
  out = rep(NA, length(filenames))

  for(i in 1:length(filenames)){
    out[i] = substr(filenames[i], attr(fs[i][[1]], 'match.length')+1, fe[[i]][1]-1)
  }
  out
}

# rollmean to find the year of passing various thresholds, then index back in to the ssp co2 concentrations
library(zoo)

findThres = function(time, x, f = 1/5, thres = 2, y.ix = 1:30){
  # smooth a timeseries and find the point where
  # it crosses a threshold.
  ysmooth = lowess(x = x, f = f)
  ysmooth.start = mean(ysmooth$y[y.ix])
  ysmooth.change = ysmooth$y - ysmooth.start
  out = time[ min(which(ysmooth.change > thres))]
  out
}

findThresRolling = function(time, y, conc, k = 20, thres = 2, bp = 1861:1880){
  # smooth a timeseries with rollmean and find the point where
  # it crosses a threshold.
  # Baseline is calculated using only the years that are in the baseline period
  # - if not all those years are there, the mean is calculated using a shorter
  # baseline period.
  
  require(zoo)
  
  # find the index with the start of the baseline period
  bp.ix = which(round(time) %in% bp)
  bp.mean = mean(y[bp.ix], na.rm = TRUE)
  
  ysmooth = rollmean(y, k = k, na.pad = TRUE)
  #NonNAindex = which(!is.na(ysmooth))
  #firstNonNA = min(NonNAindex)
  #ysmooth.change = ysmooth - ysmooth[firstNonNA]
  ysmooth.change = ysmooth - bp.mean
  y.change = y - bp.mean
  #out = time[ min(which(ysmooth.change > thres))]
  thresyear = tryCatch(time[ min(which(ysmooth.change > thres))],
                     error=function(err) NA)
  
  thresconc = tryCatch(conc[ min(which(ysmooth.change > thres))],
                       error=function(err) NA)
  
  return(list(thresyear=thresyear, thresconc = thresconc, ysmooth = ysmooth,
              ysmooth.change = ysmooth.change,
              y.change = y.change))
}

```



### Use historical global tas compared with HadCRUT4.6 observations to weight the models.
Projections are using SSP585 (for maximum signal). There are all sorts of issues ignored here (for example, putting the maesurements on a common grid), but that shouldn't matter for a simple demonstration of relative weighting.

```{r}
# data directories
fn.ssp585 = dir('/data/users/hadaw/cmip6/areaavg/tas/', pattern = 'ssp585_r1i1p1')
fn.historical = dir('/data/users/hadaw/cmip6/areaavg/tas/', pattern = 'historical_r1i1p1')

# model name lists
mods.ssp585 = extractModelnames(fn.ssp585, startstring = 'tas_Amon_', stopstring = "_ssp585")
mods.historical = extractModelnames(fn.historical, startstring = 'tas_Amon_',
  stopstring = "_historical")

# Find models common to historical and future projection
common.mods = intersect(mods.ssp585, mods.historical)

nmods = length(common.mods)

fn.ssp585.kept = fn.ssp585[match(common.mods, mods.ssp585)]
fn.historical.kept = fn.historical[ match(common.mods, mods.historical)]

# useful
# https://stackoverflow.com/questions/17215789/extract-a-substring-in-r-according-to-a-pattern
# start with strsplit, pmatch, charmatch
```

### Load and process historical model data
Uses a time series of the last 165 years of the historical runs.

```{r fig.height = 8, fig.width = 8}

# Historical years
hyrs = 1850:2014
nyr = length(hyrs)

#variable matrix (turn this into a function?
hmat = matrix(nrow = length(fn.historical.kept), ncol = nyr)

for(i in 1:length(fn.historical.kept)){
# Open and extract data from the netcdf file
  fn = fn.historical.kept[i]
  fnpath = paste0("/data/users/hadaw/cmip6/areaavg/tas/", fn)

  nc = nc_open(fnpath)

  v = ncvar_get(nc, 'tas')
  nc_close(nc)
  
  vtail = tail(v, nyr)
  hmat[i, ] = vtail
}


par(las = 1)
matplot(hyrs, t(hmat), 
        type = 'l', lty = 'solid', main = 'raw model tas', ylab = 'tas (K)')

# Amomalize model runs to the mean of each run
hmat.means = apply(hmat, 1, mean)

# varmat.anom is the variable timeseries matrix anomaly
hmat.anom = sweep(hmat, 1, STATS = hmat.means)
```

## Extract future projection data
Start with a single realisation from each model (r1i1p1), to keep everything the same.  
Hadley models use f2 - they don't have f1
```{r fig.width = 7, fig.height = 7}

fyrs = 2015:2099
fnyr = length(fyrs)
futuremat = matrix(nrow = length(fn.ssp585.kept), ncol = fnyr)

for(i in 1:length(fn.ssp585.kept)){
# Open and extract data from the netcdf file
  fn = fn.ssp585.kept[i]
  fnpath = paste0("/data/users/hadaw/cmip6/areaavg/tas/", fn)

  nc = nc_open(fnpath)

  v = ncvar_get(nc, 'tas')
  #print(length(v))
  nc_close(nc)
  
  vhead = head(v, fnyr)
 
  futuremat[i, ] = vhead
}


# Amomalize to the mean
# varmat.anom is the variable timeseries matrix anomaly
futuremat.changemat = sweep(futuremat, 1, STATS = futuremat[,1])
future.diff = futuremat.changemat [,ncol(futuremat.changemat )]

```

```{r, fig.width = 10, fig.height=10}

full_trajectory <- cbind(hmat, futuremat)
full_years <- c(hyrs,fyrs)

matplot(full_years, t(full_trajectory), type = 'l', lty = 'solid')


```

## Anomalize to early years
```{r}
anom_years <- 1881:1900
anom_index <- which(full_years%in%anom_years)
historical_mean <- apply(full_trajectory[, anom_index], 1, FUN = mean)

full_trajectory_anomaly = sweep(full_trajectory, 1, STATS = historical_mean)

par(las = 1)
matplot(full_years, t(full_trajectory_anomaly),
        type = 'l', lty = 'solid', col = 'darkgrey',
        bty = 'n',
        main = 'tas anomaly',
        ylab = 'Deg C')

```

```{r}
# Looks like we can get the ssp concentration data from here
# http://greenhousegases.science.unimelb.edu.au/#!/ghg?scenarioid=9&mode=downloads

co2_hist <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_CMIP_UoM-CMIP-1-2-0_gr1-GMNHSH_0000-2014.csv')

plot(co2_hist$year, co2_hist$data_mean_global, type = 'l')


co2_ssp119 <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_ScenarioMIP_UoM-IMAGE-ssp119-1-2-1_gr1-GMNHSH_2015-2500.csv')

co2_ssp126 <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_ScenarioMIP_UoM-IMAGE-ssp126-1-2-1_gr1-GMNHSH_2015-2500.csv')

co2_ssp245 <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_ScenarioMIP_UoM-MESSAGE-GLOBIOM-ssp245-1-2-1_gr1-GMNHSH_2015-2500.csv')

co2_ssp370 <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_ScenarioMIP_UoM-AIM-ssp370-1-2-1_gr1-GMNHSH_2015-2500.csv')

co2_ssp434 <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_ScenarioMIP_UoM-GCAM4-ssp434-1-2-1_gr1-GMNHSH_2015-2500.csv')

co2_ssp460 <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_ScenarioMIP_UoM-GCAM4-ssp460-1-2-1_gr1-GMNHSH_2015-2500.csv')

co2_ssp534 <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_ScenarioMIP_UoM-REMIND-MAGPIE-ssp534-over-1-2-1_gr1-GMNHSH_2015-2500.csv')

co2_ssp585 <- read.csv('../data/conc/mole-fraction-of-carbon-dioxide-in-air_input4MIPs_GHGConcentrations_ScenarioMIP_UoM-REMIND-MAGPIE-ssp585-1-2-1_gr1-GMNHSH_2015-2500.csv')

```


# sort the co2 data into the correct format.
```{r}
process_co2_data <- function(hist_co2, ssp_co2, years){

  # could get rid of the row names here - they just count rows from 1
  
  co2_hist_ssp <- as.data.frame(rbind(cbind(hist_co2$year, hist_co2$data_mean_global),
                         cbind(ssp_co2$year, ssp_co2$data_mean_global)))
  
  colnames(co2_hist_ssp) <- c('year','data_mean_global')
  
  out <- co2_hist_ssp[which(co2_hist_ssp$year%in%years), ]
  out
}


# could use assign() here


co2_ssp119_const <- process_co2_data(hist_co2 = co2_hist,
                           ssp_co2 = co2_ssp119,
                           years = full_years)

co2_ssp126_const <- process_co2_data(hist_co2 = co2_hist,
                           ssp_co2 = co2_ssp126,
                           years = full_years)
co2_ssp245_const <- process_co2_data(hist_co2 = co2_hist,
                           ssp_co2 = co2_ssp245,
                           years = full_years)
co2_ssp370_const <- process_co2_data(hist_co2 = co2_hist,
                           ssp_co2 = co2_ssp370,
                           years = full_years)

co2_ssp434_const <- process_co2_data(hist_co2 = co2_hist,
                           ssp_co2 = co2_ssp434,
                           years = full_years)

co2_ssp460_const <- process_co2_data(hist_co2 = co2_hist,
                           ssp_co2 = co2_ssp460,
                           years = full_years)
co2_ssp534_const <- process_co2_data(hist_co2 = co2_hist,
                           ssp_co2 = co2_ssp534,
                           years = full_years)

co2_ssp585_const <- process_co2_data(hist_co2 = co2_hist,
                           ssp_co2 = co2_ssp585,
                           years = full_years)
```


```{r}
#doesmn't look like we have anything for ssp534

cbPal <- rev(c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"))#, "#CC79A7"))

ssplist <- c('119', '126', '245', '370', '434', '460', '585')

co2_list <- paste0('co2_ssp', ssplist, '_const')

# run through each of the SSPs

plot(co2_ssp585_const, type = 'n', bty = 'n')

for(i in 1:length(co2_list)){
  co2 <- get(co2_list[i])
  lines(co2, col = cbPal[i], lwd = 1.5)
}

```

```{r}
getTAShist <- function(fn.historical.kept, hyrs = 1850:2014){
  
  nyr = length(hyrs)

  #variable matrix (turn this into a function?
  hmat = matrix(nrow = length(fn.historical.kept), ncol = nyr)


  for(i in 1:length(fn.historical.kept)){
  # Open and extract data from the netcdf file
    fn = fn.historical.kept[i]
    fnpath = paste0("/data/users/hadaw/cmip6/areaavg/tas/", fn)

    nc = nc_open(fnpath)

    v = ncvar_get(nc, 'tas')
    nc_close(nc)
  
    vtail = tail(v, nyr)
    hmat[i, ] = vtail
  }
  hmat
}

```


```{r}
getTASssp <- function(ssp, mods.historical, fyrs = 2015:2099){
  
  fn.ssp <- dir('/data/users/hadaw/cmip6/areaavg/tas/', pattern = paste0('ssp',ssp,'_r1i1p1'))
  mods.ssp = extractModelnames(fn.ssp, startstring = 'tas_Amon_', stopstring = paste0('_ssp', ssp))
  
  common.mods = intersect(mods.ssp, mods.historical)
  nmods = length(common.mods)
  fn.ssp.kept = fn.ssp[match(common.mods, mods.ssp)]
  
  fnyr = length(fyrs)
  futuremat = matrix(nrow = length(fn.ssp.kept), ncol = fnyr)
  
  for(i in 1:length(fn.ssp.kept)){
    # Open and extract data from the netcdf file
    fn = fn.ssp.kept[i]
    fnpath = paste0("/data/users/hadaw/cmip6/areaavg/tas/", fn)
    
    nc = nc_open(fnpath)
    
    v = ncvar_get(nc, 'tas')
    #print(length(v))
    nc_close(nc)
    
    vhead = head(v, fnyr)
    vl <- length(vhead)
    
    futuremat[i,1:vl ] = vhead
  }
  
  return(list(tas = futuremat, mods = common.mods, fn.ssp = fn.ssp, mods = mods.ssp))
}

# interestingly, the SSPs have different year lengths: make sure we're using the right section
# (hacked for now)
```


```{r}
test_ssp <- getTASssp(ssp = '119', mods.historical = mods.historical, fyrs = 2015:2099)

matplot(t(test_ssp$tas), type = 'l')


  mods.historical = extractModelnames(fn.historical, startstring = 'tas_Amon_',
  stopstring = "_historical")

# Find models common to historical and future projection
  common.mods = intersect(test_ssp$mods, mods.historical)

  nmods = length(common.mods)

  #fn.ssp.kept = fn.ssp[match(common.mods, mods.ssp)]
  fn.historical.kept = fn.historical[ match(common.mods, mods.historical)]

test_hist <- getTAShist(fn.historical.kept = fn.historical.kept)

matplot(t(test_hist), type = 'l')


```


```{r}

# 
for(i in 1:length(ssplist)){
  print(i)
  
  fyrs <- 2015:2099
  hyrs <- 1850:2014
  TASssp <- getTASssp(ssp = ssplist[i], mods.historical = mods.historical, fyrs = 2015:2099)
  
  # find a way to match what's coming out here
  
  # data directories
  #fn.ssp = TASssp$fn.ssp
  #fn.historical = dir('/data/users/hadaw/cmip6/areaavg/tas/', pattern = 'historical_r1i1p1')

# model name lists
  mods.ssp = TASssp$mods
  
  mods.historical = extractModelnames(fn.historical, startstring = 'tas_Amon_',
  stopstring = "_historical")

# Find models common to historical and future projection
  common.mods = intersect(mods.ssp, mods.historical)

  nmods = length(common.mods)

  #fn.ssp.kept = fn.ssp[match(common.mods, mods.ssp)]
  fn.historical.kept = fn.historical[ match(common.mods, mods.historical)]
  
  TAShist <- getTAShist(fn.historical.kept)

# hmat will change each time  
  full_trajectory_ssp <- cbind(TAShist, TASssp$tas)
  
  assign(paste0('full_trajectory_tas_ssp',ssplist[i]), full_trajectory_ssp)
  
  assign(paste0('models_tas_ssp',ssplist[i]), TASssp$mods)
  
  
  
}




```


```{r}
for(i in 1:length(ssplist)){
  
  matplot(t(get(paste0('full_trajectory_tas_ssp',ssplist[i]),)), type = 'l')
  
}

```




## Year of passing 1.5 and 2 degrees for SSP585 
```{r}
# re-write as a function
passyears_ssp585 <- as.data.frame(matrix(NA, nrow = length(common.mods), ncol = 5))
colnames(passyears_ssp585) <- c('modname', "year_1.5","conc_1.5", "year_2", "conc_2")

passyears_ssp585[,1] <- common.mods 

for(i in 1:length(common.mods)){
  
  out1_5 <- findThresRolling(time = co2_ssp585_const$year, y = full_trajectory[i,],
                         conc = co2_ssp585_const$data_mean_global, k = 20, thres = 1.5, bp = 1861:1880)
  
  passyears_ssp585[i, "year_1.5"] <- out1_5$thresyear
  passyears_ssp585[i, "conc_1.5"] <- out1_5$thresconc
  
  
  out2 <- findThresRolling(time = co2_ssp585_const$year, y = full_trajectory[i,],
                         conc = co2_ssp585_const$data_mean_global, k = 20, thres = 2, bp = 1861:1880)
  
  passyears_ssp585[i,"year_2"] <- out2$thresyear
  passyears_ssp585[i,"conc_2"] <- out2$thresconc
}


generatePassyears <- function(mods, co2, tas, k = 20, bp = 1861:1880){
  
  # mods is a vector of model names, length the same as rows in tas
  
  stopifnot(length(mods)==nrow(tas))
  
  passyears <- as.data.frame(matrix(NA, nrow = length(mods), ncol = 5))
  
  colnames(passyears) <- c('modname', "year_1.5","conc_1.5", "year_2", "conc_2")
  
  passyears[,1] <- mods 
  
  for(i in 1:length(mods)){
    
    out1_5 <- findThresRolling(time = co2$year, y = tas[i,],
                               conc = co2$data_mean_global, k = k, thres = 1.5, bp = bp)
    
    passyears[i, "year_1.5"] <- out1_5$thresyear
    passyears[i, "conc_1.5"] <- out1_5$thresconc
    
    out2 <- findThresRolling(time = co2$year, y = tas[i,],
                             conc = co2$data_mean_global, k = k, thres = 2, bp = bp)
    
    passyears[i,"year_2"] <- out2$thresyear
    passyears[i,"conc_2"] <- out2$thresconc
  }
  
  passyears
  
}

#test <- generatePassyears(mods = common.mods, co2 = co2_ssp585_const, tas = full_trajectory)

```


```{r}
# need to swap out tas for the right one each time

for(i in 1:length(co2_list)){
  
  passyears <- generatePassyears(mods = get(paste0('models_tas_ssp',ssplist[i])), co2 = get(co2_list[i]), 
                                 tas = get(paste0('full_trajectory_tas_ssp', ssplist[i])))
  
  assign(paste0('passyears_ssp',ssplist[i]), passyears) 
  
}



```

```{r}
par(las = 1)
matplot(full_years, t(full_trajectory_anomaly),
        type = 'l', lty = 'solid', col = 'darkgrey',
        bty = 'n',
        main = 'SSP585',
        ylab = 'Deg C')

points(passyears_ssp585[,'year_1.5'],rep(1.5, length(common.mods)), pch = 20, col = 'skyblue2')
points(passyears_ssp585[,'year_2'],rep(2, length(common.mods)), pch = 20, col = 'tomato2')

```
```{r}

for(i in 1:length(ssplist)){
  
  anom_years <- 1881:1900
  anom_index <- which(full_years%in%anom_years)
  
  full_trajectory <- get(paste0('full_trajectory_tas_ssp',ssplist[i]))
  historical_mean <- apply(full_trajectory[, anom_index], 1, FUN = mean)
  
  full_trajectory_anomaly = sweep(full_trajectory, 1, STATS = historical_mean)
  
  par(las = 1)
  matplot(full_years, t(full_trajectory_anomaly),
        type = 'l', lty = 'solid', col = 'darkgrey',
        bty = 'n',
        main = paste0('SSP',ssplist[i]),
        ylab = 'Deg C')
  
  passyears = get(paste0('passyears_ssp',ssplist[i]))

 points(passyears[,'year_1.5'],rep(1.5, length(get(paste0('models_tas_ssp', ssplist[i])))), pch = 20, col = 'skyblue2')
  points(passyears[,'year_2'], rep(2, length(get(paste0('models_tas_ssp', ssplist[i])))), pch = 20, col = 'tomato2')
  
}



```

